---
title: "Review"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

1.  **Clone the Homework Repository**\
    Clone the GitHub repository containing the homework file to your **local machine**.

2.  **Rename the File**\
    Rename the `.qmd` file from `hw01.qmd` to: `hw01_yourusername.qmd`.\

For example: `hw01_semiyari.qmd`

### Working on Your Homework

-   As you complete each part of the homework:

    -   **Stage** and **commit** your `.qmd` file *after each question/sub-question*, such as `1.1`, `1.2`, etc.\
    -   **Render** your `.qmd` file to **HTML only** (no PDF or Word formats).\
    -   **Push** your updates to GitHub after completing each **main question**, such as `1.0`, `2.0`, etc.

> Committing and pushing to GitHub **are graded components** of this assignment.

## Final Checks Before Submission

-   **Re-render** to HTML before your final commit and push.\

-   **Check your output:**\

    -   Make sure **all images display correctly**.\
    -   Avoid including **excessive output or large datasets** unless asked.

## Final Submission

Once you have finished and pushed your final version:

-   Copy the link of your **GitHub Pages site** (where the rendered HTML is published).
-   **Post the link** to the Canvas assignment submission page.

```{r}
# Install pacman if it's not already installed
if (!require("pacman")) install.packages("pacman")

# Use pacman to install (if needed) and load the packages
pacman::p_load(pacman, tidyverse,
               nycflights13)
```

-   Consider the data in the nycflights13 package.

-   **Problem 2**: Is there a relationship between the age of a plane and its delays?

    ```{r}
    library(nycflights13)
    library(dplyr)
    library(ggplot2)

    planes_age <- planes |>
      mutate(age = 2013 - year) |>
      select(tailnum, age)

    flights_with_age <- flights |>
      left_join(planes_age, by = "tailnum")

    valid_data <- flights_with_age |>
      filter(!is.na(age), !is.na(arr_delay))



    ggplot(valid_data, aes(x = age, y = arr_delay)) +
      geom_point(alpha = 0.1) +
      geom_smooth(method = "lm", se = FALSE, color = "blue") +
      labs(title = "Scatter plot of Plane Age vs Arrival Delay",
           x = "Plane Age (years)",
           y = "Arrival Delay (minutes)")

    correlation <- cor(valid_data$age, valid_data$arr_delay)

    print(correlation)


    ```

    We can not observe any patter that would indicate a relationship on the scatter plot. The correlation statistic is -0.01767153 which indicates a very small negative relationship. In conclusion there is no significant relationship.

-   **Problem 3**: Find the 10 days of the year that have the highest median departure delay, then select all flights from those 10 days.

    ```{r}
    flights$date <- as.Date(paste(flights$year, flights$month, flights$day, sep = "-"))

    temp_summary <- summarize(group_by(flights, date), median_dep_delay = median(dep_delay, na.rm = TRUE))
    temp_summary <- arrange(temp_summary, desc(median_dep_delay))
    top_10_days <- slice_head(temp_summary, n = 10)

    top_10_days

    flights_top10days <- filter(flights, date %in% top_10_days$date)

    flights_top10days
    ```
