---
title: "WebSraping-SQL"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3, fig.width = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

## Instructions

Complete the following exercises using the appropriate packages in R. Ensure that your solutions are optimized and use functional programming principles where applicable.

1.  Load the necessary libraries.
2.  Answer each question in separate R code chunks.
3.  Provide detailed explanations for your approach.
4.  Submit the rendered HTML file.

```{r}
  if (!require("pacman")) install.packages("pacman")

# Load contributed packages with pacman
pacman::p_load(pacman,rvest, dplyr, tidyverse, xml2,janitor, DBI, duckdb, nycflights13)
```

## WebScraping

### Problem 1:

Go to the following page [List of U.S. states and territories by population](https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_population) and select the table `State and territory rankings`.

```{r}
install.packages(c("rvest", "janitor", "dplyr"))  # Only if not already installed
library(rvest)
library(janitor)
library(dplyr)

```

1.  Make sure the name of columns and the first row is correct

    ```{r}

    # Load libraries

    library(rvest)

    library(dplyr)

    library(stringr)

     

    # URL of the Wikipedia page

    url <- "https://en.wikipedia.org/wiki/List_of_U.S._states_and_territories_by_population"

     

    # Read the webpage and extract the table 

    webpage <- read_html(url)

    tables <- html_table(webpage, fill = TRUE)

    state_table <- tables[[1]]
    ```

    \
    2.Make sure the class of each column is correct

    ```{r}

    # Fix duplicate column names first
    colnames(state_table) <- make.names(colnames(state_table), unique = TRUE)


    # Set column names accorning to the table
    colnames(state_table) <- c("State_Territory", "Population_2024_Est", "Population_2020_Census",              "Change_Percent_2010_2020", "Change_Absolute_2010_2020",
    "House_Seats", "House_Seats_Change", "Pop_Per_Electoral_Vote",              "Pop_Per_House_Seat", "Percent_of_US_2020", "Percent_of_EC_2020")

    # Ensure each data type is correct

    state_table <- state_table |>

      mutate(
    State_Territory = as.character(State_Territory),

    Population_2024_Est = as.numeric(gsub("[^0-9]", "", Population_2024_Est)),

    Population_2020_Census = as.numeric(gsub("[^0-9]", "", Population_2020_Census)),

    Change_Percent_2010_2020 = as.numeric(gsub("[^0-9.-]", "", Change_Percent_2010_2020)),

    Change_Absolute_2010_2020 = as.numeric(gsub("[^0-9.-]", "", Change_Absolute_2010_2020)),

    House_Seats = as.numeric(gsub("[^0-9]", "", House_Seats)),

    House_Seats_Change = as.numeric(gsub("[^0-9.-]", "", House_Seats_Change)),

    Pop_Per_Electoral_Vote = as.numeric(gsub("[^0-9]", "", Pop_Per_Electoral_Vote)),

    Pop_Per_House_Seat = as.numeric(gsub("[^0-9]", "", Pop_Per_House_Seat)),

    Percent_of_US_2020 = as.numeric(gsub("[^0-9.]", "", Percent_of_US_2020)),

    Percent_of_EC_2020 = as.numeric(gsub("[^0-9.]", "", Percent_of_EC_2020))
    )

     
    #final results
    print("Final column names:")
    print(colnames(state_table))
    print("\nColumn classes:")
    print(sapply(state_table, class))
    print("\nFirst few rows:")
    print(head(state_table))
    print(paste("\nTotal rows:", nrow(state_table)))
    ```

    \
    \
    Replace `__` with `NA`

    ```{r}
    library(dplyr)

    # Replace __ with NA

    state_table <- state_table |>

      mutate(across(everything(), ~ ifelse(. == "__" | . == "—" | . == "", NA, .)))
    ```

    ```         
    ```

    \

    \

## SQL

Please download a duck database. <https://data-science-master.github.io/lectures/data/flights.duckdb>

-   Use `duckdb()` and `DBI::dbConnect()` to create a connection to "flights.duckdb".

    ```{r}
    # Load required libraries
    library(DBI)
    library(duckdb)

    # Create data directory
    if (!dir.exists("data")) {
      dir.create("data")
    }

    # Download the flights database
    download.file("https://data-science-master.github.io/lectures/data/flights.duckdb", destfile = "./data/flights.duckdb", mode = "wb")

    ```

```{r}

     practcon<- dbConnect(duckdb(dbdir = "./data/flights.duckdb", read_only = FALSE))

```

-   `dbConnect()`: This function establishes a connection to a database.

-   `duckdb()`: Specifies you are connecting to a DuckDB database, which is an embedded analytical SQL database.

-   `dbdir = "flights.duckdb"`: This tells DuckDB to store the database in a file called flights.duckdb. If the file doesn't exist, it will be created.

-   `read_only = FALSE`: This means you are allowing the database to be modified. If you set this to TRUE, the database would be read-only, meaning you could only query data but not make changes to it.

-   So, setting `read_only = FALSE` is necessary if you want to add, update, or delete data in the database.

-   You can see the object `practcon` in the "Global Environment" indicating the connection profile has been established to the database.

-   Check is the connection has established

```{r}
library(DBI)


dbIsValid(practcon)
```

Please note that a basic SQL code chunk looks like this (put SQL code between the chunks):

```{r, echo = FALSE, comment = ""}
    codechunk <- "```{sql, connection=con}\n\n```"
    writeLines(codechunk)
```

Print out a summary of the tables in this database.

```{sql, connection=practcon, echo=FALSE, eval=FALSE}
    DESCRIBE;
```

-   Answer the following question using SQL.

### Problem 2: Select Flights from a Specific Carrier

Write a SQL query to find all flights from the carrier "United Air Lines Inc.".

```{sql, connection=practcon}

-- company list
SELECT DISTINCT carrier
FROM flights;

-- selecting United Airlines(UA) flights
SELECT *
FROM flights
WHERE carrier = 'UA';



```

### Problem 3: Count the Number of Flights for Each Carrier

Write a SQL query to count the total number of flights for each carrier.

```{sql, connection=practcon}

SELECT carrier, COUNT(*) AS total_flights
FROM flights
GROUP BY carrier



```

\

## Remember to Close the connection When You Are Done

-   if the connection is open then you need to use `dbDisconnect()` function

    ```{r}

    ```

## Submission

-   Save your Quarto file after completing each problem, render it, then stage and commit the changes.\

-   Include well-commented code and clear explanations.\

-   Make at least **one commit per question**.\

-   Submit the rendered file to **Canvas** and push it to your **GitHub repository**.\

-   Provide the link to your GitHub repository in the **Canvas gradebook** as a text submission.\

🔹 **Problem 1 is worth 40 points, and Problems 2 and 3 are each worth 20 points.**\

🔹 You will receive an additional **20 points** if you correctly push your completed file to GitHub and make at least one commit for each problem.\

**Good luck!**
